Thermodynamic Structure of the Solar Corona: Tomographic Reconstructions and MHD Modeling

### 2020-01-15 lloveras_2020.tex
version mergeada con conclusiones

\begin{abstract}
Observational {techniques} play an essential role in advancing our understanding of the physics of the solar corona. They provide validation data {for} three-dimensional (3D) magnetohydrodynamic (MHD) models of the solar atmosphere, {key to improve} their space weather forecasting capabilities. Solar rotational tomography (SRT) is currently the sole observational technique that provides an empirical 3D description of some {of the fundamental} plasma parameters of the solar corona at a global scale. Based on EUV data of space borne instruments, SRT allows constructing 3D maps of the coronal electron density and temperature at heliocentric heights below 1.25 Rsun. We carry out a study of the corona combining tomographic reconstructions with MHD simulations {using} the latest version of the Alfv\'en Wave Solar Model (AWSoM) of the Space Weather Modeling Framework (SWMF). Target rotations were selected from the solar minimum between solar cycles (SC) 23 and 24 and the current declining phase of solar cycle 24. Tomographic reconstructions and results of the model are {analysed} in distinct coronal magnetic structures. We study magnetically closed structures associated with the equatorial streamer belt, as well as magnetically open regions enclosing it. We report on the tomographic results in the different structures, their implications for the physics of the solar corona, and the {current} capability of the AWSoM model to reproduce the tomographic reconstructions in different regions.

\end{abstract}

@@

@@

{
\begin{itemize}
\item Type 0: closed-small-down with footpoints in the range $|{\rm latitude}|<50\mdeg$. 
\item Type I: closed-small-up with footpoints in the range $|{\rm latitude}|<50\mdeg$. 
\item Type II: closed-large-up with footpoints in the range $|{\rm latitude}|>40\mdeg$. 
\item Type III: open with footpoint in the range $|{\rm latitude}|>60\mdeg$.
\end{itemize}
}

In Section \ref{resu}, the results of both the DEMT and AWSoM models in the distinct magnetic strucures are statistically analysed. {As shown in Section \ref{demt_res}, classification of legs from type 0 through II allows studying the streamer belt in three distinct, progressively outer layers.}

@@

{For both target rotations, the top panels of Figure \ref{rpoint_demt} show the latitude-longitude location (at heliocentric height $r=1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. Open legs are indicated in gray color and closed ones in black color. For each leg, the fits to tomografic temperature and density were applied, as given by Equations (\ref{Nfit})-(\ref{Tfit}). Considering the DEMT data points and the resulting fits along each leg, the bottom panels of Figure \ref{rpoint_demt} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a four-color code, type 0, I, II and III legs are shown in blue, red, {magenta and cyan} color, respectively.} Of the {$\approx 48000$} legs selected for CR-2082, 21\% are type 0, 31\% are type I, 22\% are type II and 26\% type III. On the other hand, of the {$\approx 57000$} legs selected for CR-2208, 9\% are type 0, 36\% are type I, 26\% are type II and 29\% type III.

@@

{Throughout the magnetically closed region of both rotations, type 0, I and II legs, associated to  increasingly outer regions of the equatorial streamer belt, progressively exhibit decreasing coronal base density, increasing density scale height, and increasing electron temperature. In both rotations also, type III legs in the CHs are characterized by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I legs in the core of the equatorial streamer.}

@@

\notebyalbert{We could quantify the anti-correlation between the AWSoM terminal speed and the DEMT electron density at the coronal base in the following fashion. We can set up starting points at, say, $20\,\mrsun$, every one degree, both in latitude and longitude. This would mean setting up $180\times 360$ starting points. We trace then the (open) field lines from each of those starting points down to $1.0\,\mrsun$ to find out the DEMT results at those lower heights for each field line. Now, to do this we need to develop some codes and have no time for this paper. We can postpone it for a next effort, or... Chip: do you think you could do this with your present tools? Do you think it is interesting? I think so. Our submission deadline is end-of-January.}

@@

\notebyalbert{Propongo que Diego escriba una primer sección limpia, ordenada. Lo que vos mandarías al Solar Physics, sin asteriscos ni nada raro. Expresate con párrafos limpios. Si querés destacar algo para discutir ponelo como una nota en rojo al final del párrafo correspondiente, no en el medio de los mismos. Yo creo que esto debe estructurarse así: 
\begin{enumerate}
\item Un párrafo (breve) de que se hizo, esto contextualiza. 
\item Dos o tres párrafos sobre los resultados 3.1. Esto es: resultados per se de cada rotación y como comparan (mencionando el tema incertezas). Estructura 3D del streamer en CAPAS, revelada observacionalmente en forma global por DEMT, esto incluye UP/DOWN loops. Explicar que con AIA no andamos bien en CHs. O sí? Energía.
\item Uno o dos párrafos sobre los resultados 3.2. Como compara AWSoM con DEMT. En forma breve. Vale la pena explicar que DOWN loops no se puede.
\item Luego un párrafo sobre next steps. Esto puede incluir nuevas cosas con AIA-4, el Eclipse, otras rotaciones de mayor intensidad.
\end{enumerate}
}

\notebyalbert{Este párrafo que sigue es de Albert. Algo así debe ir incluído dentro de la parte que refiere a 3.1. Creo que hay que corregir números a la luz de los análisis finales. Creo que los números de cuanto más frío o caliente, etc, deben ser corregidos a la luz de la nueva Tabla 1.}

In this work the combination of DEMT reconstructions and PFSS models has been used to perform a comparative study of the 3D thermodynamical state of the low solar corona ($1.02-1.23\,\mrsun$) {for two rotations sampling} the last two solar minima. Specifically the study compares rotations CR-1915 and CR-2081, {respectively sampling} the SC 22/23 and SC 23/24 minima. The SC 22/23 minimum was shorter, more active and intermittent. The SC 23/24 minimum {was remarkably} characterized by a $\approx 1$ year period {showing virtually} no sunspots. Being CR-2081 near the end of that extremely quiet period, it was one of the most axisymmetric and quiet rotations on record from that period \citep{nuevo_2013}.

In comparing the DEMT results obtained for the two selected targets, it is important to bear in mind they rely on data provided by two different instruments: AIA and EUVI. In order to quantify the systematic difference of the DEMT products based on both instruments, \citet{nuevo_2015}, who were the first to apply DEMT to AIA data, {analysed} a single target using both instruments independently. They found that while the density product is essentially equal, the temperature product of DEMT based on AIA data is systematically 8\% larger than the one based on EUVI data, i.e. $\Tm^{\rm(AIA)}/\Tm^{\rm(EUVI)} \approx 1.08$. Considering such correction, Figure \ref{histos_fulldemt} and Table \ref{tabla_demt} indicate that CR-2208 was $\approx 10-15\%$ hotter than CR-2082 throughout the streamer belt region. As for the electron density products, CR-2208 was found to be $\approx 15-20\%$ less dense than CR-2082 throughout the streamer belt region. These systematic differences are beyond the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} estimated to be $\lesssim 5\%$ and $\approx 2\%$ for the electron temperature and density, respectively. \notebyalbert{son estos errores compatibles con los informados en la Sección 2?}

\notebyalbert{El texto hasta el final son de Diego. Algunos se pueden ir, otros pueden servir de semilla para construir las distintas partes de la Sección.}

*We found down legs in the DEMT reconstructions mainly in the equatorial latitudes, being a greater quantity and better distribution along longitudes in CR-2082 than in CR-2208. \citet{nuevo_2013} showed a decrease in the amount of down loops as the solar activity increases, which is consistent with the results shown in this article.

*We can see, on type 0 loops, a median value of energy input flux ($\phi_h$) for CR-2082 twice bigger than CR-2208. In both rotations this magnitude has negative values, this can refer to some loops where loop-integrated radiative flux $\phi_r$) is too small in comparison with the larger gradients of negative conductive flux $\phi_c$), resulting in negative energy input flux.

This effect can be due a subestimation of the density present in the loop or the limited range of temperature used to reconstruct tomographic parameters, that can affect the compute of radiative power. Nevertheless, this population represent lower than the \diego{ALGÚN PORCENTAJE$\%$} of total population in both cases.

Consistent that the density in CR-2082 was higher, the $\phi_r$ was also higher. In both rotations the magnitudes in the temperature gradients are similar, but the temperatures in CR-2208 were consistently higher at all heights, resulting in a larger $\phi_c$ in module. This results are in agree with the observations of \citet{maccormack_2017}.

\diego{THE AR of CR-2082 wasn't well modeled by AWSoM.}

In both rotations, AWSoM and DEMT resutls shows same termodynamic structures along latitude and longitude with very comparable values between $1.055 - 1.25\,\mrsun$

Figure \ref{histos_2082} and Table \ref{tabla_comp} indicate that AWSoM reconstruction of CR-2082 was found to be less dense $\approx 11-22\%$ and colder $\approx 5-10\%$.

Both reconstructions of CR-2208 reconstructed the ARs with very similare values and in almost the same latitude-longitude location. Figure \ref{histos_2208} and Table \ref{tabla_comp} indicate that AWSoM reconstruction of CR-2208 was found to be denser $\approx 0-10\%$ and colder $\approx 10-15\%$.

CR-2208 was $\approx 10-15\%$ hotter than CR-2082 throughout the streamer belt region. As for the electron density products, CR-2208 was found to be $\approx 15-20\%$ less dense than CR-2082 throughout the streamer belt region. These systematic differences are beyond the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} estimated to be $\lesssim 5\%$ and $\approx 2\%$ for the electron temperature and density, respectively.

\diego{I would like to say something about FIgure 10. Maybe the weird tale in AWSoM Ne in the poles. Ideas?}\\

%\diego{We could say something about the anti correlation between speed and density and not get into the trace because issues do always appear and we don't have much time, and we all want to be happy in january. Unless the collaborators find easy to make this part.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Acknowledgements
%
% \begin{acks}
%
% \end{acks}

>>>>

\begin{abstract}
Observational techniques play an essential role in advancing our understanding of the physics of the solar corona. They provide validation data for three-dimensional (3D) magnetohydrodynamic (MHD) models of the solar atmosphere, key to improve their space weather forecasting capabilities. Solar rotational tomography (SRT) is currently the sole observational technique that provides an empirical 3D description of some {of the fundamental} plasma parameters of the solar corona at a global scale. Based on EUV data of space borne instruments, {differential emision measure tomography (DEMT)} allows constructing 3D maps of the coronal electron density and temperature at heliocentric heights below 1.25 Rsun. We carry out a study of the corona combining {DEMT} reconstructions with MHD simulations using the latest version of the Alfv\'en Wave Solar Model (AWSoM) of the Space Weather Modeling Framework (SWMF). Target rotations were selected from the solar minimum between solar cycles (SC) 23 and 24 and {the declining} phase of SC 24. The 3D tomographic and MHD models are analysed and compared in distinct coronal magnetic structures. {We report in quantitative detail on the 3D termodinamic structure of the inner corona as revealed by the DEMT analysis. We show in particular the existence of diverse types of temperature structures in the low and mid latitudes of the streamer belt, and that the heating energy required to keep them stable may be accounted for by Alfvén waves. We show that the latest version of the AWSoM model results agree with DEMT reconstructions to within a $\approx 20\%$ accuracy or better. We report on aspects of the AWSoM model that could be improved to better reproduce the DEMT results, in particular concerning the source region of the fast/slow solar winds.}
\end{abstract}

@@

In the present work, data from 3 EUV bands was used. 171, 193 and 211 \AA \ in case of AIA, and 171, 195 and 284 \AA \ bands in case of EUVI. When using data from 3 EUV bands, a Gaussian model for the LDEM is able to accurately predict the tomographic emissivities \citep{frazin_2009,nuevo_2015}. In each tomographic voxel the problem is then reduced to finding the values of the three free parameters of the Gaussian (centroid, standard deviation, and area) that allow to best reproduce the three tomographically reconstructed values of FBE in that voxel.

@@

{
\begin{itemize}
\item Type 0: closed-small-down with footpoints in the range $|{\rm latitude}|<50\mdeg$. 
\item Type I: closed-small-up with footpoints in the range $|{\rm latitude}|<50\mdeg$. 
\item Type II: closed-large-up with footpoints in the range $|{\rm latitude}|>40\mdeg$. 
\item Type III: open with footpoint in the range $|{\rm latitude}|>60\mdeg$.
\end{itemize}
}
{Legs of type down turn out to have a significant population only for closed-small loops. In the case of closed-large loops and open field lines, there are very few cases of type down. Hence, the requirement of being up for legs of type II and III is included above to select the vastly dominating population in each case}. \\
In Section \ref{resu}, the results of both the DEMT and AWSoM models in the distinct magnetic strucures are statistically analysed. {As shown in Section \ref{demt_res}, classification of legs from type 0 through II allows studying the streamer belt in three distinct, progressively outer layers.}

@@

{For both target rotations, the top panels of Figure \ref{rpoint_demt} show the latitude-longitude location (at heliocentric height $r=1.105\,\mrsun$) of all traced field line legs for which criterion (i) of Section \ref{trace} is met. Open legs are indicated in gray color and closed ones in black color. For each leg, the fits to tomografic temperature and density were applied, as given by Equations (\ref{Nfit})-(\ref{Tfit}). Considering the DEMT data points and the resulting fits along each leg, the bottom panels of Figure \ref{rpoint_demt} show the latitude-longitude location of the subset for which also both criteria (ii) and (iii) of Section \ref{trace} are met. Using a four-color code, type 0, I, II and III legs are shown in blue, red, {magenta and cyan} color, respectively.} Of the {$\approx 44000$} legs selected for CR-2082, 20\% are type 0, 31\% are type I, 23\% are type II and 26\% type III. On the other hand, of the {$\approx 44000$} legs selected for CR-2208, 10\% are type 0, 43\% are type I, 30\% are type II and 17\% type III.

@@

{Throughout the magnetically closed region of both rotations, type 0, I and II legs, associated to  increasingly outer regions of the equatorial streamer belt, exhibit progressively decreasing coronal base density, increasing density scale height, and increasing electron temperature. In both rotations also, type III legs in the CHs are characterized by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I legs in the core of the equatorial streamer.}

@@

\notebyalbert{We could quantify the anti-correlation between the AWSoM terminal speed and the DEMT electron density at the coronal base in the following fashion. We can set up starting points at, say, $20\,\mrsun$, every one degree, both in latitude and longitude. This would mean setting up $180\times 360$ starting points. We trace then the (open) field lines from each of those starting points down to $1.0\,\mrsun$ to find out the DEMT results at those lower heights for each field line. Now, to do this we need to develop some codes and have no time for this paper. We can postpone it for a next effort, or... Chip: do you think you could do this with your present tools? Do you think it is interesting? I think so. Our submission deadline is 24 January.}

@@

%\notebyalbert{Propongo que Diego escriba una primer sección limpia, ordenada. Lo que vos mandarías al Solar Physics, sin asteriscos ni nada raro. Expresate con párrafos limpios. Si querés destacar algo para discutir ponelo como una nota en rojo al final del párrafo correspondiente, no en el medio de los mismos. Yo creo que esto debe estructurarse así: 
%\begin{enumerate}
%\item Un párrafo (breve) de que se hizo, esto contextualiza. 
%\item Dos o tres párrafos sobre los resultados 3.1. Esto es: resultados per se de cada rotación y como comparan (mencionando el tema incertezas). Estructura 3D del streamer en CAPAS, revelada observacionalmente en forma global por DEMT, esto incluye UP/DOWN loops. Explicar que con AIA no andamos bien en CHs. O sí? Energía.
%\item Uno o dos párrafos sobre los resultados 3.2. Como compara AWSoM con DEMT. En forma breve. Vale la pena explicar que DOWN loops no se puede.
%\item Luego un párrafo sobre next steps. Esto puede incluir nuevas cosas con AIA-4, el Eclipse, otras rotaciones de mayor intensidad.
%\end{enumerate}
%}

{We carried out the 3D DEMT reconstruction and MHD modeling of the inner solar corona for two target rotations, CR-2082 and CR-2208. The first target belongs to the deep minimum epoch between SCs 23 and 24, while the latter belongs to the end of the declining phase of SC 24. The present work introduces two improvements in the implementation of DEMT, namely the use of only off-limb data in the EUV images, and the inclusion of a 3D regularization scheme. As a result of the former, artifacts introduced by coronal dynamics are mitigated, and as a result of the latter, tomographic reconstructions behave more smoothly close to the radial boundaries of the computational grid when compared to previous reconstructions. The MHD model used is here the latest version of the AWSoM component of the SWMF suite.}

{Based on the magnetic field model of AWSoM model and the results of the DEMT model for electron density and temperature, we classified coronal structures in four different types. Smaller legs in the core of the streamer belt are classified in down/up types according to their temperature decreasing/increasing with height (dubbed type 0/I). Larger closed legs of type up that form the envelope of the streamer belt (dubbed type II), and open field lines of type up populating the CHs (dubbed type III).}

{Concerning the DEMT reconstructions, throughout the magnetically closed region of both rotations, we find that type 0, I and II legs are associated to increasingly outer regions of the equatorial streamer belt, and exhibit progressively decreasing coronal base density, increasing density scale height, and increasing electron temperature. For both target rotations we find that down legs populate the low latitudes of the streamer belt, while up legs dominate its mid-latitudes. Also, in the case of CR-2082 the fraction of down legs is significantly larger than for CR-2208. These findings are consistent with previous studies by \citet{huang_2012} and \citet{nuevo_2013}. In the case of the latter, they include in their analysis target CR-2081, which is a rotation almost identical to our target CR-2082. Our results for target CR-2082 compare very well with those of \citet{nuevo_2015} and \citet{lloveras_2017} for target CR-2081. As our study uses the improved version of the DEMT technique, such comparison provided a consistency check for the changes implemented in the DEMT technique.}

{In both rotations also, type III field lines in the CHs are characterized by sub-MK temperatures, and electron density values of order $\approx 1/2$ of those observed for the type 0 and type I lines in the core of the equatorial streamer. In the case of the AIA based DEMT reconstruction, we show that in a substantial fraction of the CHs region the technique performs poorly in modeling LDEMs that predict the tomographic emissivitiy in each EUV band with reasonable accuracy. As a result, DEMT results based on AIA data are currently less reliable in the CH region. This was already found by the existing two previous DEMT works based on AIA data \citep{nuevo_2015,maccormack_2017}. The source of this distinct performance of DEMT when applied to EUVI or AIA data is related to the TRFs of the respective filter sets. We are currently implementing an improved LDEM determination algorithm to to deal with this issue that will be used in a follow up article.}

In comparing the DEMT results obtained for the two selected targets, {we highlight here that} they rely on data provided by two different instruments, {namely EUVI and AIA for CR-2082 and CR-2208, respectively.} In order to quantify the systematic difference of the DEMT products based on both instruments, \citet{nuevo_2015}, who were the first to apply DEMT to AIA data, {analysed} a single target using both instruments independently. They found that while the density {products are} essentially equal, the temperature product of DEMT based on AIA data is systematically 8\% larger than the one based on EUVI data, i.e. $\Tm^{\rm(AIA)}/\Tm^{\rm(EUVI)} \approx 1.08$. Considering such correction, Figure \ref{histos_fulldemt} and Table \ref{tabla_demt} indicate that CR-2208 was {characterized by temperatures $\approx 10-15\%$ larger relative to CR-2082} throughout the streamer belt region. As for the electron density products, CR-2208 was found to be $\approx 15-20\%$ less dense than CR-2082 throughout the streamer belt region. These systematic differences are {around or beyond the uncertainty level in the DEMT products due to systematic sources (radiometric calibration and tomographic regularization), that \citet{lloveras_2017} {estimated to be $\Delta\Tm\approx 10\%$ and $\Delta\sqravgN\approx 5\%$}.}

{In the streamer belt region we combined the DEMT results and the AWSoM magnetic model to infer the energy input flux $\phi_h$ at the coronal base required to maintain stable coronal loops. To this end, we applied the technique developed by \citet{maccormack_2017}. We found characteristic values in the range $\phi_h\approx 0.5-1.5 \times 10^5\,\erg\,\cminvs\,\s^{-1}$, depending on the rotation and the type of {loop, matching the values reported by} \citet{maccormack_2017}. Based on spectroscopic data of the EIS instrument in quiet-Sun regions \citet{hahn_2014} showed that, if the observed non-thermal broadening are assigned to Alfvén waves, their energy flux at the coronal base is estimated to be in the range $\approx 1.5-2.5\times 10^5\,\erg\,\cminvs\,\s^{-1}$. A large fraction of the coronal base energy input flux $\phi_h$ estimated in this work, or even its totality, could then be accounted for by Alfvén waves.}

{Results of the AWSoM model are compared to those of the DEMT reconstructions in the range of coronal heights $\approx 1.05-1.20\,\mrsun$ simultaneously covered by both models. In its current implementation, the AWSoM model can not reproduce down loops (type 0), so the comparison is restricted to type I, II and III field lines. The detailed comparison shown in this work can be summarized as follows. For CR-2082, the electron density of both models agree within $\approx 20\%$ in all regions, while for CR-2208 the agreement is within $\approx 5\%$. In the case of the electron temperature, both models agree within $\approx 10-15\%$ in all regions, for both target rotations. This level of agreement between both models is considerably better than that reported in previous works. \citet{oran_2015} and \citet{jin_2012} reported electron density values of the AWSoM model typically 50\% larger than those of the DEMT reconstructions.} \notebyalbert{Chip, if Section 2.2 explains the most significant improvements of AWSoM since the version used in those references, we can comment on the reason for the improvement here.}

{Within the quantitative match summarized in the previous paragraph, there are some qualitative agreements and disagreements between both models worth noticing. For both target rotations, the AWSoM model reproduces the relatively lower temperatures found by DEMT to characterize the low-latitudes of the equatorial streamer belt compared to its mid-latitudes. On the other hand, while the latitude of the open/close magnetic boundary in both hemispheres matches the location of the strongest latitudinal gradient of the DEMT electron density, this is not the case for the AWSoM model, that shows a minimum density at the open/closed boundary. Also, while the DEMT electron density decreases from the open/closed boundary towards the poles (in both hemispheres of the two target rotations), the AWSoM model shows the opposite trend. This behavior is notoriously opposite to that reported in the AWSoM model version used by \citet{oran_2015}, in which the electron density decreases from the open/closed boundary towards the poles.} \notebyalbert{Chip: Even if the match AWSoM/DEMT is quantitatively pretty good, this qualitative opposite behavior is off putting. Do you think there is room for improvement in the AWSoM model here? Maybe worth mentioning?}

\notebyalbert{Chip: In relation to the previous comment, if the anlaysis proposed in Figure 15 is carried out we should say something about that here.}

{A detailed empirical description of the 3D termodynamic structure of the inner corona at a global scale is curently only possible with tomographic techniques, such as DEMT. Using tomographic results for continuous validation of 3D MHD models is of high relevance for the continued improvement of models. In a follow up article we will analyse whit DEMT and 3D MHD modeling new target rotations of the current solar minimum between SCs 24 and 25, in particular the rotations conrresponding to the 2019 and 2020 total solar eclipses.}

\noindent
\notebyalbert{Diego, agregá las sección de agradecimientos, fijándote como lo hicimos para tu paper anterior. Vos seguro debes agradecer a la beca, Ceci tambien (fijate en su paper), y todos los de aqui deberemos agraecer grants. Armalo como puedas y luego lo revisamos.}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Acknowledgements
%
 \begin{acks}
D.G.Ll. and C.M.C. acknowledges CONICET doctoral fellowship (Res. Nr. 4870) to IAFE that supported his participation in this research.
The authors acknowledge \diego{ANPCyT grant 2012/0973 and CONICET grant PIP \# 11220120100403 estos son viejos?} to IAFE that partially supported their participation in this research
 \end{acks}
